---
title: "Develop Data Products Assignment"
author: "gtargus"
date: "October 30, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Betting Analyzer

Betting analyzer is a very basic app to compute your potential profits and give you a also a possibility of success  

Obviously this is an app meant for fun :-)

## The formula

###If we want to win 100$ by betting on 2 games then the decimal odd is 10 (10X10=100)  
###In general:  
####*Decimal Odd= exponential ((log ( "desired earning" / "number of predictions "))*  

###Decimal odd is connected to  Win % (or Break Even %) with this formula:  

###Decimal odd:X+1 / Fractional odd:X /Break Even %:(100/X+1)%

## ui.R side
```{r, eval=FALSE,echo=TRUE}
library(shiny)
shinyUI(fluidPage(
 titlePanel("Betting Analyzer"),
  sidebarLayout(
    sidebarPanel(
      h3("Give me amount of money "),
sliderInput("money","Earnings in dollars:",min = 100,max = 2000000,value = 1000),
h3("Give me number of games "),sliderInput("num_bets","Num of games:",min = 1,max = 15,value = 5)
    ),
    mainPanel(
      h3("The average per game return to make that cash is :"),
      textOutput("bettingAnalysis"),
      h3("The probability of making that cash is(e-4)% :"),
      textOutput("Odds")
    )
  )
))
```

## server.R side
```{r, eval=FALSE,echo=TRUE}
library(shiny)

shinyServer(function(input, output) {

  bettingAnalysis<- reactive ({exp((log(input$money)/input$num_bets))  })
  num_bets<- reactive ({input$num_bets })   

  Odds<-reactive ({1/(bettingAnalysis() ^ num_bets()) *10000})  
output$bettingAnalysis=renderText(bettingAnalysis())
  output$Odds=renderText(Odds())
})

```
